---
title: "Homework3Quarto"
format: html
editor: visual
---

## Task 1: Conceptual Questions

**1. If your working directory is myfolder/homework/, what relative path would you specify to get the file located at myfolder/MyData.csv?** \> You would specify ../MyData.csv.

**2. What are the major benefits of using R projects?** \> Some of the major benefits of using R projects are the following: 1. R projects follow an organized structure 2. Each R project has its own working director to prevent conflicts and 3. R projects are easily reproducible.

**3. What is git and what is github?** \> Git tracks changes while GitHub hosts repositories. GitHub is within Git.

**4. What are the two main differences between a tibble and a data.frame?** \> Tibbles print a limited number of rows while data.frame prints the rows and columns in full. The tibble function always returns a tibble while a data.frame can be manipulated to produce vectors and/or matrices as well as entire data frames.

**5. Rewrite the nested function listed in the pdf homework 3 instructions using baseR's chaining operator.** \> iris %\>% as_tibble() %\>% select(starts_with("Petal"), Species) %\>% filter(Petal.Length \< 1.55) %\>% arrange(Species)

**6. What is meant by long format data and wide format data? Which do we generally prefer for statistical analysis?** \> In long format data, each row represents a single observation. In wide format data, each row represents a single entity. We generally prefer long format data for statistical analysis.

## Task 2: Reading Delimited Data

### Glass Data

```{r}
# Install packages
install.packages("tidyverse")
install.packages("DBI")
install.packages("RSQLite")
```

```{r}
# Load packages
library(tidyverse)
library(readxl)
library(dplyr)
library(readr)
library(DBI)
#library(RSQLite)
```

#### 1. Read in the data
```{r}
# Read the data directly from the URL
url1 <- "https://www4.stat.ncsu.edu/~online/datasets/glass.data"

# Add the column names
glass <- read_csv(url1, col_names = c("Id", "RI", "Na", "Mg", "Al", "Si", "K", "Ca", "Ba", "Fe", "Type_of_glass"))

# Print the tibble
glass
```

#### 2. Start a chain to overwrite using mutate()

```{r}
# Start a chain to overwrite the Type_of_glass variable using mutate()
glass <- glass %>%
  mutate(Type_of_glass = if_else(Type_of_glass==1, "building_windows_float_processed",
                                 if_else(Type_of_glass==2, "building_windows_non_float_processed",
                                         if_else(Type_of_glass==3, "vehicle_windows_float_processed",
                                                 if_else(Type_of_glass==4, "vehicle_windows_non_float_processed",
                                                         if_else(Type_of_glass==5, "containers",
                                                                 if_else(Type_of_glass==6, "tableware", "headlamps")))))))
```

#### 3. Continue your chain and keep only specific observations

```{r}
# Keep only observations where the Fe variable is less than 0.2 and where the type of glass is either "tableware" or "headlamp"

glass2 <- glass %>%
  mutate(Type_of_glass = if_else(Type_of_glass==1, "building_windows_float_processed",
                                 if_else(Type_of_glass==2, "building_windows_non_float_processed",
                                         if_else(Type_of_glass==3, "vehicle_windows_float_processed",
                                                 if_else(Type_of_glass==4, "vehicle_windows_non_float_processed",
                                                         if_else(Type_of_glass==5, "containers",
                                                                 if_else(Type_of_glass==6, "tableware", "headlamps"))))))) %>%
  filter(Fe < 0.2 & (Type_of_glass=="tableware" | Type_of_glass=="headlamps"))
```

### Yeast Data

#### 1. Read in the data

```{r}
# Read the data in from the URL
url2 <- "https://www4.stat.ncsu.edu/~online/datasets/yeast.data"

# Assign the data to an object
yeast <- read_delim(url2, delim = "  ", col_names = FALSE)

# Add column names
colnames(yeast) <- c("seq_name", "mcg", "gvh", "alm", "mit", "erl", "pox", "vac", "nuc", "class")

# Print the tibble
yeast
```

#### 2. Start a chain that removes specific columns

```{r}
yeast2 <- yeast %>%
  select(-seq_name) %>%
  select(-nuc)
```

#### 3. Continue your chain to add corresponding columns

```{r}
yeast3 <- yeast %>%
  select(-seq_name) %>%
  select(-nuc) %>%
  group_by(class) %>%
  mutate(across(c(mcg, gvh, alm, mit, erl, pox, vac), list(mean = mean, median = median), .names = "{.col}_{.fn}")) %>%
  ungroup()
```

## Task 3: Combining Excel and Delimited Data

#### 1. Read in the excel version of the white wine data set using readxl

```{r}
# Import the data from the first sheet
excel1 <- "C:/Users/natal/Desktop/NCSU/summer 24/ST558/Repos/Homework3/white-wine.xlsx"

# Read the first sheet of the file
whitewine <- read_excel(excel1)

# Print out the tibble
whitewine
```

#### 2. Rename the variables

```{r}
# Read in the variables names from the second sheet
col_names <- read_excel(excel1, sheet = 2, range = "A1:A13")
col_names <- col_names[[1]]

# Overwrite the current column names
colnames(whitewine) <- col_names

whitewine
```

#### 3. Add a column to indicate the wines are white

```{r}
whitewine <- whitewine %>%
  mutate(type = 'white')
```

#### 4. Read in the excel version of the red wine data set using readr

```{r}
# Read this in using the readr package
csv1 <- "C:/Users/natal/Desktop/NCSU/summer 24/ST558/Repos/Homework3/red-wine.csv"

redwine <- read_delim(csv1, delim = ";")

# Replace the variable names
col_names <- read_excel(excel1, sheet = 2, range = "A1:A13")
col_names <- col_names[[1]]
colnames(redwine) <- col_names

# Append a column denoting the type as "red"
redwine <- redwine %>%
  mutate(type = 'red')
```

#### 5. Combine these two data sets

```{r}
combined_wine <- bind_rows(whitewine, redwine)
```

#### 6. Filter the data

```{r}
filtered_wine <- combined_wine %>%
  filter(quality > 6.5, alcohol < 132)
```

#### 7. Continue your chain to now sort the data

```{r}
sorted_wine <- filtered_wine %>%
  arrange(desc(quality))

sorted_wine
```

#### 8. Continue your chain to select only specific variables

```{r}
selected_wine <- sorted_wine %>%
  select(contains("acid"), alcohol, type, quality)
```

#### 9. Continue your chain to add mean and standard deviation

```{r}
new_wine <- selected_wine %>%
  group_by(quality) %>%
  summarize(mean_alcohol = mean(alcohol), sd_alcohol = sd(alcohol))
```

## Task 4: Database Practice

#### 1. Connect to the database and then look at all of the tables in the database

```{r}
lahman_db <- "C:/Users/natal/Desktop/NCSU/summer 24/ST558/Repos/Homework3/lahman.db"

con <- dbConnect(RSQLite::SQLite(), dbname = lahman_db)

```

#### 2. Return all of the data from the Teams table for the year 2015

```{r}
# Construct the tabl
Teams <- tbl(con, "Teams")

# Filter for year 2015
Teams_2015 <- Teams %>%
  filter(year == 2015) %>%
  collect()

# Return all of the specified data
Teams_2015
```

#### 3. Repeat the above with the sql() function within tbl()

```{r}
Teams_2015 <- tbl(con, sql("SELECT * FROM Teams WHERE year = 2015"))

Teams_2015
```

#### 4. Return all of the hall of fame players

```{r}
hof <- tbl(con, "HallOfFame")

hof_players <- hof %>%
  filter(inducted == "yes") %>%
  select(playerID, yearID, category)

hof_players
```

#### 5. Combine the table

```{r}
ppl <- tbl(con, "People")

combined_table <- hof %>%
  filter(inducted == "yes") %>%
  left_join(ppl, by = "playerID") %>%
  select(nameFirst, nameLast, yearID, category)

combined_table
```

#### 6. Return only the specified columns

```{r}
# Return only the playerID, G, W, and L columns from the Managers table
Managers <- tbl(con, "Managers")

filtered_managers <- Managers %>%
  select(playerID, G, W, L)

# Determine the overall win/loss records for each of these HOF managers and sort
win_loss <- filtered_managers %>%
  group_by(playerID) %>%
  summarize(total_games = sum(G, na.rm = TRUE),
            total_wins = sum(W, na.rm = TRUE),
            total_losses = sum(L, na.rm = TRUE),
            w_l = total_wins/total_losses,
            Total_W/G_managed = total_wins/total_games) %>%
  arrange(desc(Total_W/G_managed))
```

#### 7. Return information for only the people that managed a team and were inducted into the HOF

```{r}
hof_managers_join <- inner_join(Managers, hof, by = "playerID")

hof_managers <- hof_managers_join %>%
  filter(inducted == "yes")

hof_managers
```
